---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%", 
  eval = FALSE
)
```

## martini

<!-- badges: start -->
[![R-CMD-check](https://github.com/Bayer-Group/bmdi-mlai-martini/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/Bayer-Group/bmdi-mlai-martini/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

`{martini}` provides a convenient framework to collect data from multiple 
clinical trial data sets following the ADaM standard and create a 
machine-learning-ready data set.  

It is part of a set of packages that form a pipeline for machine learning and 
artificial intelligence applications in clinical research that offers a 
streamlined and standardized way to assess relationships between clinical study 
data and a specified outcome while aligning with validation and documentation 
standards expected for software used in clinical trials.  

While designed to interface smoothly with internal downstream modeling and 
reporting tools, the outputs from `{martini}` can of course also be used with any 
ML framework of your choice.

The core of `{martini}` revolves around three key steps, each represent by one 
main function:  

- `spec()` – Specify the relevant steps to collect and combine the relevant data  
- `build()` – Build one raw analysis dataset in wide format with the relevant clinical features  
- `prepare_ml()` – Prepare a clean, analysis-ready data object for modeling  


## Installation

To install the latest version of `martini` from GitHub:

```{r, eval=FALSE}
remotes::install_git(
  "https://github.com/Bayer-Group/bmdi-mlai-martini",
  dependencies    = TRUE,
  build_vignettes = TRUE
)
```


## Documentation

Please refer to the package vignettes for an exemplary analysis either by 
running the following code or by opening the  
[pkgdown site](https://bayer-group.github.io/bmdi-mlai-martini/).

```{r eval = FALSE}
vignette(package = "martini")
```
