---
title: "MARTINI - MLAI pipeline prepare module"
subtitle: "Document data preparation for TLF"
author: "Maike Ahrens, Sebastian Voss"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{MARTINI - TLF}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
library(martini)
library(tidyselect)
library(dplyr)
library(recipes)
library(purrr)
library(gt)
library(magrittr)

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval    = FALSE
)
```

# Scope

This vignette shows the CSR output for the example of a regression outcome, 
included in the {martini} package:

* `martini_spec` the general specification
* `martini_ml_regr` the output of `prepare_ml`, containing the `recipe` object,
the ML data set and more

Questions to be answered:

**A How was the ML model built?** 

* definition of feature space 
* list of preprocessing steps with main settings
* tuned hyper parameters

**B Characterisation of ML model**

* most important variables (VIP table and plot)
* model performance

**C Is the ML model applicable to new data**

* all variables used in the ML model available in new study?
* descriptive summary of the variables 
  * numerics: to check mean/sd/ranges/units/skewness for numerics incl correlation structure
  * factors: compare level sets for factors
  
**D Preprocessing of new data**

* if applicable: transform to match units from training, recode levels
* `bake()`
   * proportion missing per variable pass/fail `step_filter_missing()`
   * missing observations in non-removed columns for subjects with non-missing outcome missing values 
   * list of removed subjects and columns
   
   
**E Results**

* scatter subject predictions and actual outcome
* Characterization in terms of trt estimate (location, CI) and (adjusted) r^2
  * reference model
  * model with added ML score


# A model building

## Feature space

The `raw` feature space is defined as all variables technically
available from the selected data sets after applying 
the user-defined filters.

This can be derived from the output of `prepare_ml`.

```{r}
colnames_raw <- martini::martini_ml_regr %>% 
  get_data('raw') %>% 
  names() %>% 
  setdiff(c('.id', '.out'))

martini::martini_ml_regr$dict %>% 
  filter(param %in% colnames_raw) %>% 
  select(
    column, 
    param, 
    Label = label, 
    Domain = spec_id
  ) %>% 
  gt()
 # TODO column/label usage for occds
```




```{r, eval = FALSE}
#This can be derived from the martini `spec` object. 

dict_raw <- martini_spec %>% 
  map('dict') %>% 
  list_rbind(names_to = 'dataset') 

dict_raw %>% 
  filter(selected) %>% 
  select(dataset, param, label) %>%
  filter(!param %in% martini_spec$adsl[c('id')]) %>% 
  DT::datatable()
```


The actual feature space available to the ML methods will most likely differ, 
where there may be additional columns, column removals and renaming, respectively.


```{r}
martini_spec$adlb[c('id', 'trt')]

```


```{r}
martini_ml_regr$dict
```






* manual selections
* spec

# recipe object

List of pre-processing operations with general setting and 
the list of variables 
that the recipe step is trained for

```{r}
preprocess_steps(martini_ml_regr$prep_recipe) %>% 
  kable() %>% 
  kable_styling(full_width = FALSE)
```



